Generative Simplex Mapping: Nonlinear Endmember Extraction and Unmixing for Hyperspectral Imagery

### 2024-07-16 paper/paper.tex
added some equations

@@

@@

>>>>

\newpage

@@

\begin{equation}\label{eqn:data-space-distribution}
    p(x \mid z, \mathbf{W}, \beta) = \left(\frac{\beta}{2\pi} \right)^{D/2}\exp\left( -\frac{\beta}{2}\lVert \psi(z; \mathbf{W}) - x \rVert^2 \right)
\end{equation}

\begin{equation}\label{eqn:latent-distribution}
    p(z) = \sum\limits_k^K \pi_k \, \delta(z - z_k) \quad\text{where}\quad  \sum_k^K\pi_k = 1
\end{equation}

\begin{equation}\label{eqn:llh}
    \mathcal{L}(\mathbf{W}, \beta) = \sum\limits_n^N \ln \left( \sum\limits_k^K \pi_k \, p(x_n \mid z_k, \mathbf{W}, \beta) \right)
\end{equation}

The activation function $\mathbf{\Phi}_{km} = \phi_m(z_k)$ is chosen to enable modeling linear and nonlinear mixing. For $m \leq N_v$, we take $\Phi_{km} = z_k$. For $m > N_v$, we set
\begin{equation}\label{eq:act-function}
    \mathbf{\Phi}_{km} = \begin{cases}
        \frac{s - \lVert z_k - \mu_m \rVert}{s}; & \lVert z_k - \mu_m \rVert \leq s \\
        0; & \lVert z_k - \mu_m \rVert > s
    \end{cases}
\end{equation}
The activation function centers $\mu_m$ are selected to be greater than

We include a prior distribution on model weights to enable
\begin{align}\label{eq:weight-prior}
\begin{split}
    p(\mathbf{W} \mid \lambda_e, \lambda_w) &= \left(\frac{\lambda_e}{2\pi}\right)^{\frac{D\,N_v}{2}}\exp\left(-\frac{\lambda_e}{2}\sum\limits_d^D\sum\limits_{m=1}^{N_v}\mathbf{W}_{dm}^2 \right)\\ 
    & \qquad \cdot \left(\frac{\lambda_w}{2\pi}\right)^{\frac{D(M-N_v)}{2}}\exp\left(-\frac{\lambda_w}{2}\sum\limits_d^D\sum_{m=N_v+1}^M \mathbf{W}_{dm}^2\right)
\end{split}
\end{align}

\begin{equation}\label{eq:responsibility}
    R_{kn}  = p(z_k \mid x_n, \mathbf{W}, \beta) = \dfrac{p(x_n \mid z_k, \mathbf{W}, \beta)}{\sum\limits_{k'}^K p(x_n \mid z_{k'}, \mathbf{W}, \beta)}
\end{equation}

Penalized complete-data log likelihood
\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}\label{eq:complete-data-llh}
    Q = \sum\limits_n^N\sum\limits_k^K R_{kn} \left(\ln\pi_k + \frac{D}{2}\ln\left(\frac{\beta}{2\pi}\right) - \frac{\beta}{2}\sum\limits_d^D\left(\sum\limits_m^M \mathbf{W}_{dm}\mathbf{\Phi}_{km} - \mathbf{X}_{nd}\right)^2\right) + \frac{MD}{2}\ln\left(\frac{\lambda}{2\pi}\right) - \frac{\lambda}{2}\sum\limits_d^D \sum\limits_m^M \mathbf{W}_{dm}^2
\end{equation}
\end{adjustwidth}

\begin{equation}\label{eq:pi-update}
    \pi_k^{\text{new}}  = \frac{1}{N}\sum_n R_{kn}
\end{equation}

\begin{equation}\label{eq:beta-update}
    \frac{1}{\beta^{\text{new}}}  = \frac{1}{ND}\sum\limits_n^N\sum\limits_k^K R_{kn}\lVert \psi(z_k; \mathbf{W}) - x_n \rVert^2
\end{equation}

@@

The mixing coefficients are initially chosen so that $\pi_k = \frac{1}{K}$.
